---
title: "Perturbation"
author: "Euan_McDonnell"
date: "22 May 2020"
output: html_document
---


---
title: "GeneMeta"
author: "Euan_McDonnell"
date: "21 May 2020"
output: html_document
---



# Load Libraries and biomaRt ID conversion file

```{r}
library(OmnipathR)
library(igraph)
library("crossmeta")
library("biomaRt")

ensembl <- useMart("ensembl",dataset="hsapiens_gene_ensembl")
biomart <- getBM(attributes=c('uniprot_gn_id','ensembl_gene_id'),mart = ensembl)[-1,]
colnames(biomart) <- c("ENSEMBLID","UNIPROTID")

```



# Read in files from folder

Data is read in as a list of data.frames each from the same experiment (GSEXXXXX) with each dataframe containg the Ensembl Gene IDs and Perturbation Scores. 

```{r}

setwd("######## YOUR DIRECTORY ##########"
ReadInForGraph <- function(gse_id)
{
  file.match <- paste(gse_id,".*.perturbation_time_series.",sep="")
  readin <- lapply(list.files(pattern=file.match), function(file)
  {
    read.table(file,header=TRUE,row.names=1)
  })
}

gse37827.g <- ReadInForGraph("GSE37827") # SARS CALU3 DATASET
gse56677.g <- ReadInForGraph("GSE56677") # MERS CALU3 DATASET

```



# This converts IDs from Ensembl gene ids to Uniprot gene id

```{r}


Convert2Uniprot <- function(by_time)
{
  up <- lapply(by_time, function(x)
  {
    x <- merge(biomart,x,by="ENSEMBLID")
  })
}

up.gse37827 <- Convert2Uniprot(by_time.37827)
up.gse56677 <- Convert2Uniprot(by_time.56677)

```




# Load in total PPI interaction data-sets from OmnipathR

Use the OmnipathR API to download PPI interaction data and extract from this dataframe the genes present in each time-series specific dataframes. Converts these interaction dataframes to igraph objects

```{r}

interactions <- import_Omnipath_Interactions(filter_databases=
                      c("ARN","CA1","CancerCellMap","DeathDomain","Guide2Pharma","Macrophage",
                        "NRF2ome","PDZbase","SignaLink3","Signor","SPIKE","TRIP","BioGRID",
                        "DIP","HPRD","InnateDB","IntAct","MatrixDB","MPPI","Vidal HI-III"))


GetDirectedGraph <- function(data)
{
  graphs <- lapply(data, function(time)
  {
    colnames(time)[3] <- "Pscore"
    subset <- interactions[interactions$source %in% time$UNIPROTID & 
                         interactions$target %in% time$UNIPROTID, ]
    if (nrow(time) > 1)
    {
      OPI_g <- interaction_graph(interactions = subset)
      OPI_g <- simplify(OPI_g, remove.multiple = TRUE, remove.loops = TRUE,edge.attr.comb = igraph_opt("edge.attr.comb"))
      vertex_attr(OPI_g, "Pscore") <- time$Pscore
      OPI_g
    }
  })
  names(graphs) <- names(data)
  graphs <- graphs[!sapply(graphs, is.null)]
  graphs
}
 

gse37827.g <- GetDirectedGraph(up.gse37827)
gse56677.g <- GetDirectedGraph(up.gse56677)


```




# Plot Degree distribution & Networks

Some functions to plot the degree distribution of networks and the networks themselves. 


```{r}

plot_degree_distribution <- function(graph,name) 
{
  d = degree(graph, mode = "all")
  dd = degree.distribution(graph, mode = "all", cumulative = FALSE)
  degree = 1:max(d)
  probability = dd[-1]*length(d)
  nonzero.position = which(probability != 0)
  probability = probability[nonzero.position]
  degree = degree[nonzero.position]
  if (length(probability) > 0 & length(degree) > 0)
  {
    plot(probability ~ degree, log = "xy", xlab = "Degree (log)", ylab = "Number of Nodes",
       col = 1, main = name)
  }
}

# https://igraph.org/c/doc/igraph-Community.html#idm383434440480 <- module detection algs
SaveDegreeDistPlots <- function(data)
{
  lapply(seq(1,length(data)), function(i)
  {
    title <- paste("Degree Distribution :- ",names(data)[[i]],"rs",sep="")
    png(filename=paste("degree_distribution.",".",names(data)[[i]],"rs",".png",sep=""))
    plot_degree_distribution(data[[i]],title)
    dev.off()
  })
}


SaveDegreeDistPlots(gse37827.g)
SaveDegreeDistPlots(gse56677.g)


PlotGraphWithCluster <- function(g)
{
  if ( length(V(g)) > 0 & length(E(g)) > 0)
  {
    g <- delete.vertices(g,degree(g)==0)
    g <- simplify(g, remove.multiple = T, remove.loops = T, edge.attr.comb=c(weight="sum", type="ignore") )
    g <- as.undirected(g,mode="collapse",edge.attr.comb = igraph_opt("edge.attr.comb"))
    c <- cluster_louvain(g)
    l <- layout_with_fr(g)
    c_keep_ids <- as.numeric(names(sizes(c)[sizes(c) >= 5]))
    c_keep_v_idxs <- which(c$membership %in% c_keep_ids)
    g_sub <- induced_subgraph(g, V(g)[c_keep_v_idxs])
    c_sub <- c
    c_sub$names <- c$names[c_keep_v_idxs]
    c_sub$membership <- c$membership[c_keep_v_idxs]
    c_sub$vcount <- length(c_sub$names)
    c_sub$modularity <- modularity(g_sub, c_sub$membership, E(g_sub)$weight)
    plot(c_sub, g_sub,col = membership(c)[c_keep_v_idxs],layout = l[c_keep_v_idxs, ],
         mark.border = rainbow(length(communities(c)), alpha = 1)[c_keep_ids],
         mark.col = rainbow(length(communities(c)), alpha = 0.3)[c_keep_ids],vertex.label = NA,vertex.size = 5)
  }
}


lapply(gse37827.g, function(to_plot) PlotGraphWithCluster(to_plot))
lapply(gse56677.g, function(to_plot) PlotGraphWithCluster(to_plot))


```










